\documentclass[officiallayout]{tktla} 
%\documentclass[officiallayout,a4frame]{tktla}
\usepackage[latin1]{inputenc}
\usepackage{latexsym}
\usepackage{graphicx}

\title{Deep Learning Algorithms \\ for Control}
\author{Yuan Gao}
\authorcontact{gaoyuankidult@gmail.com\par
  http://www.cs.helsinki.fi/u/yuangao/}
\pubtime{September}{2015}
\reportno{0}
\isbnpaperback{000-00-0000-0}
\isbnpdf{000-00-0000-0}
\issn{1238-8645}
\printhouse{Unigrafia}
\pubpages{7} % --- remember to update this!
% For monographs, the number of the last page of the list of references
% For article-based theses, the number of the last page of the list of
% references of the preamble part + the total number of the pages of
% the original articles and interleaf pages.
\supervisorlist{Dorota Glowacka, University of Helsinki, Finland \newline  Leo K{\"a}rkk{\"a}inen, Nokia Research Center, Finland \newline Honkala Mikko Nokia Research Center}
\preexaminera{}
\preexaminerb{}
\opponent{}
\custos{}
\generalterms{}
\additionalkeywords{}
\crcshort{A.0, C.0.0}
\crclong{
\item[A.0] Example Category
\item[C.0.0] Another Example
}
\permissionnotice{
  To be presented in \ldots{} text of a long permission notice. Text of
  a long permission notice. Text of a long permission notice. Text of
  a long permission notice. Text of a long permission notice. Text of
  a long permission notice.
}

\newtheorem{theorem}{Theorem}[chapter]
\newenvironment{proof}{\noindent\textbf{Proof.} }{$\Box$}

\begin{document}

\frontmatter

\maketitle

\begin{abstract}
Controlling a complicated mechanical system to perform a certain task, for example, making robot to dance, is a traditional problem studied in the area of control theory. Many successful applications like Google BigDog\cite{Raibert2008} and Google Self-driving(#TODO cite) car have been made in accordance to the new theories found in this field.

However more evidences show that with in-cooperating with machine learning techniques in robotics can enable people to get rid of tedious engineering works of adjusting environmental parameters(#TODO cite). Many researchers like Jan Peters(#TODO cite), Sethu Vijayakumar, Stefan Schaal, Andrew Ng and Sebastian Thrun are the early explorers in this field. Based on the Partial Observable Markov Decision Process(POMDP) reinforcement learning, they contributed theory and practical implementation of several bench marks in this field.

Recently, one sub-field of machine learning called deep learning gained a lot of attention as a method attempting to model high-level abstractions by using model architectures composed by multiple non-linear layers. (for example \cite{Krizhevsky2012}). Several architectures of deep learning networks like deep belief network \cite{Hinton2006}, deep Boltzman machine \cite{Salakhutdinov2009}, convolutional neural network \cite{Krizhevsky2012} and deep de-noising auto-encoder \cite{Vincent2010} have shown its advantages in specific areas. Especially, convolutional neural network, which was invented by Krizhevsky, outperformed all the traditional feature-based machine learning techniques in ImageNet competition.

The main works of deep learning is more related to perception which deals with the problems like Sensor Fusion{OConnor2013}, Nature Language Processing(NLP)\cite{Cho2014} and Object Recognition\cite{Lenz2013}\cite{Hoffman2014}. Although considered briefly in Jürgen Schmidhuber's team\cite{Mayer2006}, the other area of robotics, namely control, remains more-or-less unexplored in the realm of deep learning.

There are two reasons about why these areas remain unexplored,

The main focus of this thesis is to introduce general learning methods for robot control problem with an emphasize on deep learning method. As a consequence, this thesis tries to describe the transitional learning method of robot as well as the emerging deep learning method.

Ultimately, the author hope the readers of this thesis, even without much background in machine learning or robotics can understand the gists of robot learning problems.
\end{abstract}

\begin{acknowledgements}
  This is a sample sentence that should look like normal text, and
  this is another. This is a sample sentence that should look like
  normal text, and this is another. This is a sample sentence that
  should look like normal text, and this is another.
\end{acknowledgements}

\tableofcontents

\mainmatter

\chapter{Reinforcement Learning}
\section{Markov Decision Process}
\section{Partially Observable Markov Decision Process}
\section{Dynamic Programming}
\section{Reinforcement Learning Methods}
\subsection{Temporal Difference Learning}
\subsection{Q-Learning}
\subsection{Adaptive Heuristic Critic}
\subsection{Prioritised Sweeping}
\subsection{Policy Gradient Methods}

\section{Classification of the Regarded RL Problems}
\subsection{High-Dimensionality}
\subsection{Partial-Observability}
\subsection{Continuous State and Action Spaces}
\subsection{Data-Efficiency}

\chapter{Recurrent Neural Networks}
\section{Feedforward Neural Networks}
\section{Recurrent Neural Networks}
\subsection{Finite Unfolding in Time}
\subsection{Overshooting}
\subsection{Dynamical Consistency}
\section{Universal Approximation}
\subsection{Approximation by FFNN}

\subsection{Approximation by RNN}
\section{Training of RNN}
\subsection{Shared Weight Extended Backpropagation}
\subsection{Learning Methods}
\subsection{Learning Long-Term Dependencies}

\section{Improved Model-Building with RNN} 
\subsection{Handling Data Noise} 
\subsection{Handling the Uncertainty of the Initial State}
\subsection{Optimal Weight Initialisation}

\chapter{Prior Arts of Combining RNN and RL}
\section{Neural Actor-Critic(idasi's group)}
\section{LSTM with POMDP objective function}
\section{PhD thesis, by Remi Coulom ?}

\section{DQN?}
\section{Hybrid Approch(RL with RNN)}
\section{Recurrent Models of Visual Attention?}       
\section{stanley gecco021 2002?}
\chapter{Experiment}
\section{RNN(LSTM) Implementation}
\section{Cart-pole Balancing Simulator}
\section{Learning a task of stacking wooden blocks}

This is a sample sentence that should look like normal text, and this
is another. This is a sample sentence that should look like normal
text, and this is another. This is a sample sentence that should look
like normal text, and this is another.


\begin{theorem}
This is a sample sentence that should look like normal text,
and this is another:
\[ y = x+3 \]
\end{theorem}

\begin{proof}
This is a sample sentence.
\end{proof}

\bibliography{library}
\bibliographystyle{alpha}

\end{document}
